Merging Datasets Through Deep learning

### 2018-09-02 paper2018/datasets.tex
cutting to size

\section{Benchmarks}
\label{datasets}
Our benchmarks were derived from Wikidata.  Specifically, we used the \textit{also known as} property from Wikidata to get alternate forms for the same entity name for people as well as companies.  For company names, we augmented the names and surface forms in Wikidata with data from the SEC \footnote{\url{https://www.sec.gov/dera/data/financial-statement-and-notes-data-set.html}}, which has former and more recent names for companies.  There were 213106 names for people from the specific dump we extracted, and 70946 names for companies.  The extracted files and the cleansing code are available on our repository. The extracted files however contained significant noise that we cleaned up programmatically.  We describe the clean up rules for people and for companies separately because they were somewhat different.  In the case of people's names, we also augmented the data so the system could learn some common rules that define variants of a person's name.  This was not possible for company names.

@@

\subsection{Augmenting people's names}
In many cases, we had no alternate forms for a person's name even if we did have their main name.  We augmented the data with the following rules.  If the main name for the person in Wikidata had two parts, we created new source forms as follows: 
\begin{itemize}
\item Last Name, First Name
\item First Name Initial . Last Name
\item Last Name , First Name Initial
\end{itemize}
If the main name in Wikidata had three parts, we created these additional source forms in addition to the ones listed above:
\begin{itemize}
\item First Name Middle Name Initial Last Name
\item Last Name, First Name Middle Name
\item Last Name, First Name Middle Initial .
\end{itemize}

>>>>

\section{Benchmarks}
\label{datasets}
Our benchmarks were derived from Wikidata.  Specifically, we used the \textit{also known as} property from Wikidata to get alternate forms for the same entity name for people as well as companies.  For company names, we augmented the names and surface forms in Wikidata with data from the SEC \footnote{\url{https://www.sec.gov/dera/data/financial-statement-and-notes-data-set.html}}, which has former and more recent names for companies.  There were 213106 names for people from the specific dump we extracted, and 70946 names for companies.  The extracted files and the cleansing code are available on our repository. The extracted files however contained significant noise that we cleaned up programmatically.  We describe the cleansing rules for people and for companies separately because they were somewhat different.  In the case of people's names, we also augmented the data so the system could learn some common rules that define variants of a person's name.  This was not possible for company names.

@@

\subsection{Augmenting people's names}
In many cases, we had no alternate forms for a person's name even if we did have their main name.  We augmented the data with the following rules.  If the main name for the person in Wikidata had two parts, we created new source forms as follows: (a) Last Name, First Name, (b) First Name Initial . Last Name, and (c) Last Name , First Name Initial.

If the main name in Wikidata had three parts, we created these additional source forms in addition to the ones listed above: (a) First Name Middle Name Initial Last Name (b) Last Name, First Name Middle Name (c) Last Name, First Name Middle Initial .

### 2018-09-02 paper2018/joins.tex
cleaning up
cutting to size

\section{Merging datasets with siamese networks}
\label{joins}

Section~\ref{join_system} describes our overall system for using siamese networks for joins.  Section~\ref{join_complexity} provides a complexity analysis for this approach.

\section{Applying deep learning models to joins}
\label{join_system}
\begin{figure}
\includegraphics[width=1.0\linewidth]{joins}
\caption{Applying deep learning models to joins}
\label{join_fig}
\end{figure}

Figure~\ref{join_fig} provides an overview of how deep learning models can be used for merging datasets.  Panel (a) in the figure are the two datasets to be joined.  The datasets are labeled arbitrarily as a left table and a right table.  For each cell value in the two columns to be joined, we obtain vector embeddings from the siamese network that was used to estimate distance for same and different surface forms for the same entity.  Note that although the siamese network has three separate networks, each network within a triplet network is in fact identical to the other two networks because they share weights.  A single version of the network with learnt weights is used to generate the embeddings for the cell value as shown in panel (b). For the left table cell values, vector embeddings are inserted into an approximate nearest neighbors index, as shown in panel (c).  For each cell value in the right table, vector embeddings are used as `query vectors' to query the approximate nearest neighbors index as shown in panel (d).  In our context, merging the datasets would involve joining the top $k$ rows in the left table that are `closest' in distance to each cell value in the right table.  Note that the choice of $k$ clearly has a direct effect on the tradeoff between precision and recall, but this is true for any type of join algorithm that is not based on equality.

>>>>

%\section{Applying deep learning models to joins}
%\label{join_system}

%\begin{figure*}[htb]
%    \centering
%    \begin{subfigure}[t]{0.24\linewidth}
%        \centering
%        \includegraphics[width=.9\linewidth]{join1}
%        \caption{Columns to be merged}
%        \label{join1}
%    \end{subfigure}%
%    ~ 
%    \begin{subfigure}[t]{0.24\linewidth}
%        \centering 
%        \includegraphics[width=.9\linewidth]{join2}
%        \caption{Create embeddings for each cell value in each column using the siamese model}
%        \label{join2}
%    \end{subfigure}
%    ~ 
%    \begin{subfigure}[t]{0.24\linewidth}
%        \centering 
%        \includegraphics[width=.9\linewidth]{join3}
%        \caption{Index embeddings for left table's cell values in a nearest neighbors index}
%        \label{join3}
%    \end{subfigure}
%    ~ 
%    \begin{subfigure}[t]{0.24\linewidth}
%        \centering 
%        \includegraphics[width=.9\linewidth]{join4}
%        \caption{For each cell in right table, find closest neighbor(s) and merge these rows}
%        \label{join4}
%    \end{subfigure}
%    \caption{Overview of joins using deep learning}
%    \label{join_fig}
%\end{figure*}

Assuming we have deep learning models that are trained to produce the right distance estimates for alternate surface forms for an entity, the models can be used for a join as follows.  For each cell value in the two columns to be joined, obtain vector embeddings from the last layer of the network.  Note that although the siamese network has three separate networks, each network is in fact identical to the other two networks because they share weights.  For the column cell values, vector embeddings are inserted into an approximate nearest neighbors index.  For each cell value in the right column, vector embeddings are used as `query vectors' to query the approximate nearest neighbors index.  In our context, merging the datasets would involve joining the top $k$ rows in the left table that are `closest' in distance to each cell value in the right table.  Note that the choice of $k$ clearly has a direct effect on the tradeoff between precision and recall, but for most practical uses of join, $k$ is usually very small (typically 1).  This has implications on what metrics we can use to evaluate joins, as we describe in section \ref{results}.
